\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-12-04 background.tex

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in the safe subset of
Java~\cite{scj2013} amenable to formal verification and timing analysis,
while avoiding all the pitfalls of programming in 'C' style languages. A
number of derived statements also exist (e.g., \texttt{sustain},
\texttt{await}, etc) that make programming reactive systems easier. The
\textit{Model of Computation} (MoC) of a simple SystemJ program is shown
in Figure~\ref{fig:2}.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT is the \textit{Best Case
  Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in digital hardware), the numerical
annotations on the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} shows the mapping of the logical time to
the physical time. For example, the logical tick \texttt{1} starts at
the first rising edge and completes at the second rising edge of the
physical clock, whereas logical tick \texttt{4} starts at the $4^{th}$
rising edge, but finishes at the $6^{th}$ rising edge of the physical
clock. This difference can be accounted for by the Java computations
required when signal \texttt{A} is input by the environment (see
Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
Esterel and SystemJ.

>>>>

Table~\ref{tab:1} presents the SystemJ kernel statements used to program
the control flow. The data flow is programmed in the safe subset of
Java~\cite{scj2013} amenable to formal verification and timing analysis,
while avoiding all the pitfalls of programming in `C' style languages. A
number of derived statements also exist (e.g., \texttt{sustain},
\texttt{await}, etc) that make programming reactive systems easier. The
\textit{Model of Computation} (MoC) of a simple SystemJ program is shown
in Figure~\ref{fig:2}.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT is the \textit{Best Case
  Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in digital hardware), the numerical
annotations on the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} shows the mapping of the logical time to
the physical time. For example, the logical tick \texttt{1} starts at
the first rising edge and completes at the second rising edge of the
physical clock, whereas logical tick \texttt{4} starts at the $4^{th}$
rising edge, but finishes at the $6^{th}$ rising edge of the physical
clock. This difference can be accounted for by the Java computations
required when signal \texttt{A} is input by the environment (see
Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
Esterel and SystemJ.

### 2013-12-04 introduction.tex

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delay model relies on
the use of relative instead of absolute real-time, i.e., a delay in
selected time units is counted from the currently executing SystemJ
statement (3) delays allow mixing behaviors with real-time features and
others with only logical time and (4) real-time delay does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify an
  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

>>>>

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delay model relies on
the use of relative instead of absolute real-time, i.e., a delay in
selected time units is counted from the currently executing SystemJ
statement (3) delays allow mixing behaviors with real-time features and
others with only logical time and (4) real-time delay does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify an
  exact \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

### 2013-12-04 related_work.tex

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). Since in our solution we calculate the
delay in logical ticks \textit{after} allocation and scheduling, it
gives the compiler developers the chance to optimize for many different
criteria such as computation time, energy, power, etc, without worrying
about or violating the real-time delay specification.

Other works such as Quartz~\cite{glog02} and Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
statements in Esterel using \texttt{pause} and special \textit{pragmas}
(annotations), respectively.  The Quartz solution is targeted at
studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et al and our approach
is that their approach requires programmers to annotate the time taken
by 'C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.

>>>>

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). In our solution we calculate the delay
in logical ticks \textit{after} allocation and scheduling, it gives the
compiler developers the chance to optimize for many different criteria
such as computation time, energy, power, etc, without worrying about or
violating the real-time delay specification.

Other works such as Quartz~\cite{glog02} and Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
statements in Esterel using \texttt{pause} and special \textit{pragmas}
(annotations), respectively.  The Quartz solution is targeted at
studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et al. and our approach
is that their approach requires programmers to annotate the time taken
by `C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.
